---
title: "Workflow of 'x_from_power()' Using the 'power_curve' Algorithm"
output:
  html_document:
    fig.align: "center"
    toc: true
    number_sections: false
bibliography: references.bib
csl: apa.csl
---

```{r echo = FALSE}
library(DiagrammeR)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Goal

This technical appendix describes the
workflow of `x_from_power()` in
[power4mome](https://sfcheung.github.io/power4mome/)
when using the `"power_curve"` algorithm.
This algorithm belongs to a family of
algorithm called surrogate function
approximation method [see @chalmers_solving_2024
for a review].

# `x_from_power()`

The following is the workflow of `x_from_power()`,
when the algorithm `"power_curve"` is used.
In simulation-based power analysis, a number of replications
is to be done for each value of `x`, and this
can be slow when Monte Carlo or bootstrapping confidence
intervals are involved in the test in each replication.
It is not feasible,
and also not necessary, to accurately estimate the levels
of power along many values across a range of `x`, if the
goal is to find a value of `x` with estimated
power close to the target power.
Therefore, steps are taken to
balance speed and precision when finding the solution.

```{r fig-n_from_power, fig.cap = "From Power to x (Sample Size or Effect Size [Parameter Value])", echo = FALSE}
# https://mermaid.live/edit#pako:eNqtV21v2kgQ_isrnyoZibShBJr41KtsDAWSVFXSfjggshazgCOz5tbr1FyU_36zb_aapNX1epGSeGbn9ZnxzPrRibMVcTxnnWbf4i1mHH0JFxShV69QnOI8D8karcgaFylH6yRNvd-G8DMa_X4sRBjLGAVjWmw0CuHnmVjOwQehKy0VhkIOpIRczg8pQQlNeIJTJYDSZLPlG0aIJZFm2d4-XqYFqU_zLC14klFb4kBSyK-WIWXCf3S-zgoTYiMCITHz3fnCGSUgwLcElcilKGOI5K2Fc9cSEr4_B4mQcMJ2CSVSzGT1gCHYHG2SB0IbB4zkgHHuoZzwqIwYphviCpN30mTgLpxhzpMd5gR9zr4R5qG9-HfGSc6j5SGibkvE0WSSXNhQUQ3AxDVhG4JwavkDqTLqGKHQFblxhJUhFBfsgWhXkSTAzckfgA2PEq0E9fVDHxS_7lciPHHWsc-C-qzpbQgHt4RLIHLyV0FoDOBka0QZ2edtdAO_GIAuc7SHYDgDpCzDoxfUhfZNbgl9_I7Qb9LPsWGhFYhcbkWjAgxHR8Hj48IZTFCcUc5xAvogtwHzEqEPC-fpyUhaiKhCQz9A-YVbhW1KHkgKBjKIC7N4W3sZNBuolEKcHTxoM9UDdYugEyhHGd2bVglCoeyv7oucSxzREudkheCNaLiFMEoov5CI1izbRfJU25Pc2uTQ7r79T3WfNCerfq-LEoxeasRaznRH8LEJ4I8aUndkp4ZwLAo1aZYHapAnK2VNoidAMIiuGvDYlRz7rvXCx9CQiQyqmjRuGWUFb6FvCd-aMNMsB8uybqBkR2HSG_tWh0gLbZ2WfJQlkE9xov6LN0EiBJSVaGM0NCMsQf4Yu8q9bFChiHcNhEVj_OxwAWvBLyYjjQz-ZWuMg_BXmmMimsOnBzSADklonBa6K773Mk98PUUg_G0iy_rl5uvQhDMJ3OZKwLoZYFIo8wndPHMgMXCXLZTvIFvRLbAd6QqzldqmlfVBnSv9z30ymT1LYeRf3Q5rialARZ-K7GwApr44NL7Q-_cAKcVpJDhNOSnoc3iPMGSUwfiSQygla34kKDD9REo1pWDYm1kfwUiOjubuVER_jctkV-wQLXZL1ahSKodGwfGWrD7U8pciClcl01JAvxx8yw7qUpa5ukCIW4CH5PJQr7F54z1ocWmirYDWz6oqmgDf-kmNWPWsmVagge2SZvzI7Sf_f3R25TbWijX7ns8J5BZibSTres5V0bVeesfKqDJXe7wGj58hJMzUCwb476EELqYiq2rtoGx5T2LeqhRns_mbhXNDeMGopbhw3tyZa5h06vv67lgsNwzvt9V9ag5eTyaQKpah38oNq3V9rRuoq5UiBuqSpIhQXYYUMVRacGvV6kPJD45dy3vp_ItoSXQFz3fmOqHEA3UtAAL9SXLN02eD-uxTZrNUPEGoNrsihmonK2KkVqoiPqq9qYjxUXhVGedi8j1fYx9MvONGiGO_ZprYxkG1Go2Qb3aaYkyq7WQ0fLN1DCMwi8MwBmYJGEZYGQmbVieNCCe-ntGaDPRM1uTAVjIZTGZHRa1gkt8G86H4K_pOFLJCZqJDm-qJWlFqRjWimvo22_idzqr512Br_KYSP9tI88A2o4do026VD3BshRet-N_DQH3_zKuxNBJ0hcK0GeNlw-Ez9mWT69dME8ZlUI1fxbjSk7GimpldKf61GTGSnEkAnLazg-mGkxV80z6K44UDVdzBrceDR_0lu3AW9AlEccGz2wONHY-zgrQdlhWbreOtYaUAVciNGyYYMNlV3D2msyzbGRWySnjGrtVHtPyWliKO9-iUjtfrvz7tdXtnZxcXp93Td_3OWds5ON5Jp9frvL7odrsX5_3zztt-t_vUdv6WZnuvu71Ov3Pa75_2em8v3p332s6GiXx0jIADYQMoCHe8zvn5287TPyG_BaM
mermaid('
flowchart TD
  %% classDef default fill:#EEEEFF;
  %% classDef errornode fill:#FFDDDD;
  %% classDef startend fill:#DDFFDD;

  style initial fill: lightgreen
  style loop fill: lightblue
  style solution fill: lightyellow
  style exit fill: lightyellow
  style found fill: lightgreen

  ZA(["Find the x (n or es)"])
  AA[["Determine the initial values given the initial results: set_x_range()"]]
  AB("Estimate Power: power4test_by_n() or power4test_by_es()")
  AC("Merge all results: by_x_1")
  AD("Fit a power curve: power_curve() -> fit_i")
  %% ADA("Update fit_1")
  %% ADB("Update by_x_1")
  AE("Set the sequences of nreps, Rs, and xs per trial")
  %% AF("Set the sequence of Rs")
  %% AG("Set the sequence of # of ns per trial")

  BA("Start a trial")

  BB{{"CI conttains target power?"}}

  BBA("Update the interval of power levels to search")

  BC[["Determine xs to try: estimate_x_range() --> x_j"]]
  BD[["Adjust nrep based on power levels of xs: nrep_from_power() --> nrep_j"]]
  BE("Estimate power: power4test_by_n() or power4test_by_es() --> by_x_j")
  BF("Merge all results --> by_x_1")
  BG("Update the power curve: power_curve() --> fit_1")

  BH{{"Is target power inside the range of estimated power levels?"}}

  BHA("Find the candidate solution (x_out) with power closest to the target power")
  BHAA("Update x_out, power_out, nrep_out, ci_out, and by_x_out")

  BHB("Estimate the candidate x by the power curve")
  BHBA("Estiamte the power of x: power4test_by_n() or power4test_by_es()")
  BHBB("Update x_out, power_out, nrep_out, ci_out, and by_x_out")
  BHBC("Merge all results --> by_x_1")
  BHBD("Update the power curve: power_curve() --> fit_1")

  BI{{"Any CIs include the target power?"}}

  BIA("Set ci_hit to TRUE")
  BIB("Find the x (a) with CI including the target power and (b) smallest standard error")
  BIC("Update n_out, power_out, nrep_out, ci_out, and by_x_out")

  BIZ("Set ci_hit to FALSE")

  BJ{{"ci_hit TRUE?"}}

  BJA{{"nrep_out == final_nrep?"}}

  BJAA{{"At least one nrep left?"}}

  BJAAA("Next nrep, R, and xs_per_trial")

  BJZ("Maximum number of trials reached?")

  BK{{"(ci_hit) and (nrep_out == final_nrep)?"}}

  BKA("Solution found: Set to the solution: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")

  BKB("Solution not found: Set to NA: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")

  BL("Determine x estimated by the power curve (used if solution not found): power_curve() --> x_estimated")

  BM("Prepare the output (an x_from_power object)")

  ZZ[/"Return the output"/]

  ZA --> AA

  subgraph initial [Pre-Iteration Search]

  AA --> AB
  AB --> AC
  AC --> AD
  AD --> AE

  end

  AE --> BA

  subgraph loop [Trial Loop]

  BA --> BB
  BB -- Yes --> BBA --> BC
  BB -- No --> BC
  BC --> BD
  BD --> BE
  BE --> BF
  BF --> BG
  BG --> BH

  subgraph solution [Any candidate solution?]

  BH -- Yes --> BHA
  BH -- No --> BHB

  BHA --> BHAA
  BHAA --> BI

  BHB --> BHBA
  BHBA --> BHBB
  BHBB --> BHBC
  BHBC --> BHBD

  BHBD --> BI

  BI -- Yes --> BIA
  BIA --> BIB
  BIB --> BIC

  BI -- No --> BIZ

  end

  subgraph exit [Exit the Loop?]

  BIC --> BJ
  BIZ --> BJ

  BJ -- Yes --> BJA

  BJ -- No --> BJZ

  BJA -- No --> BJAA
  BJAA -- Yes --> BJAAA
  BJAA -- No --> BJZ
  BJAAA --> BJZ

  end

  BJZ -- No --> BA
  BJAA -- No --> BA

  end

  subgraph found [Solution Found?]

  BJA -- Yes --> BK

  BJZ -- Yes --> BK

  BK -- Yes --> BKA
  BK -- No --> BKB

  BKA --> BL
  BKB --> BL

  end

  BL --> BM

  BM --> ZZ

', height = 4050, width = 800)
```

In `x_from_power`, `x` can be a sample
size (`n`) or a population value (`es`,
"effect size") of the selected model
parameter.

# Technical Details

These are the basic steps of the algorithm:

- Pre-iteration search

  - Several candidate values of `x` spanning
    a range is selected, using a small
    initial number of replications
    (`nrep0`). If simulation-based methods
    such as Monte Carlo or bootstrapping
    is used, a small initial number of
    resampling (`R0`) is also used.

  - The levels of power for these levels
    are estimated.

  - A working power curve is fitted to
    these values of `x` and the estimated
    levels of power.

- The search

  - The working power curve will be used
    to identify one or more probable values
    of `x` (`x_j`) to examine.

  - The levels of power for `x_j` are
    estimated.

  - The working power curve will be updated.

  - Examine the values

    - If the number of replications is
      not yet equal to the target number
      (`final_nrep`), then the search
      will continue, increasing the number
      of replications for high precision
      if necessary.

   - If the number of replications is
     already equal to the target number
     (`final_nrep`), then the values tried
     will be examined to see whether
     a solution has been found.

## Termination Criteria

If one of the following criteria is met,
the search will end:

- A solution is found. This depends on
  the `goal`. For `ci_hit`, a solution is
  found if the confidence interval of
  the estimated power includes the target
  power. For `close_enough`, a solution is
  found if the difference between the
  estimated power and the target power
  is less than the tolerance (`tolerance`).
  The tolerance can be set
  manually by the argument `tolerance`,
  and its default is .02 for `close_enough`.

- The range of changes of `x` in the `last_k`
  iterations is less than `delta_tol`.
  For `power_curve`, it suggests that
  the search failed, and a new search using
  a different seed or different interval
  is needed.
  To change `delta_tol`, set it through
  the `control` argument (e.g.,
  `control = list(delta_tol = .02)`.
  The number of trails (*k*), default to
  3, can be changed through the
  `control` argument (e.g.,
  `control = list(last_k = 5)`).

- The maximum number of trials (iterations)
  for the search is reached.
  The default value is determined by
  the calling function, default to 10 for
  `power_curve`, and
  can be changed through the argument
  `max_trials` of `x_from_power()` and
  friends.

Note that the working power curve is
only used to assist the identification
of the values of `x`
to examine. Whether a value of `x`
(e.g., a sample size) is considered a
solution is determined solely by
the estimated power, and/or the confidence
interval. Therefore, whether the working
power curve is "correct" along a wide range
of `x` is not essential.
The working power curve serves it role
as long as it can lead to a predicted
value of `x` close enough to the solution.

## The Working Power Curve

Strictly speaking, the model to be fitted
is not the true power curve, but a model
that is "good enough" for the values of
`x` examined so far (that's why these models
are called surrogate functions).

The function `power_curve()` is used
to fit a working power curve.
Unlike some previous implementations,
one or models are tried, in sequence,
to determine the working power curve.
The current internal default models for
sample sizes are as following, tried
in this order:

- `reject ~ 1 - I(exp((a - x) / b))`, fitted by `nls()`

- `reject ~ I((x - c0)^e) / (b + I((x - c0)^e))`, fitted by `nls()`

- `reject ~ x`, fitted by a logistic regression model using `glm()`

- `reject ~ x`, fitted by a linear regression model using `lm()`.

The current internal default models for
effect sizes are as following, tried
in this order:

- `reject ~ 1 - I(exp((a - x) / b))`, fitted by `nls()`

- `reject ~ 1 - 1 / I((1 + (x / d)^a)^b)`, fitted by `nls()`

- `reject ~ 1 - exp(x / a) / I((1  + exp(x / a))^b)`, fitted by `nls()`

- `reject ~ 1 - 2 / (exp(x / d) + exp(-x / d))`, fitted by `nls()`

- `reject ~ 1 / (1 + a * exp(-b * x))`, fitted by `nls()`

- `reject ~ x`, fitted by a logistic regression model using `glm()`

- `reject ~ x`, fitted by a linear regression model using `lm()`.

The term `reject` is the estimated power
for `nls()` and `lm()`, and the significant
test results (0 for not significant and 1 for significant)
for `glm()` logistic regression model.
If appropriate, the number of replications
will be used as the weights when fitting
a model.

The `nls()` models will be fitted first,
and the model with the smallest deviance
will be used. If all models failed,
which is possible for `nls()`, then the
logistic regression model will be fitted.
If this model also failed to fit, then
the linear model will be attempted.

Although the relation between power and
`x` certainly is not linear for a wide
range of `x` (e.g., sample sizes), this
algorithm only needs a working power curve
that is a good-enough approximation for
the interval being searched. Therefore,
if the interval is narrow enough and
close to solution, a simple linear model
can also be a good approximation.

# Annotation

- `by_x_1`

  - The collection of all values tried and their results.
    It is updated whenever new value(s) is/are tried.

- `fit_1`

  - The latest power curve estimated by
    `power_curve`, using the values tried,
    stored in `by_x_1`. It is updated whenever
    `by_x_1` is updated.

- `x_j`

  - The value(s) for which power levels
    will be estimated in a trial.

- `nrep_j`

  - The number of replications to be used
    when estimate the power level for
    a value of `x`. In a trial, the
    numbers of replications can be different
    for different values, for efficiency.

- `by_x_j`

  - The results for of `power4test_by_n()`
    or `power4test_by_es()` given
    `x_j` for a trial.

- `x_out`

  - The value of `x` which is a *candidate*
    solution (e.g., with estimated power
    closest to the target value).

- `power_out`, `nrep_out`, `ci_out`, `by_x_out`

  - Results based on `x_out`.

- `ci_hit`

  - Logical. Set to `TRUE` if there is at
    least one value of `x` with the
    confidence interval of the estimated
    power including the target power.

- `final_nrep`

  - The desired number of replications
    for the solution. This value determines
    the desired level of precision
    (the width of the confidence interval)
    in the solution.

- The sequences of values for `nrep`, `R`,
  and the number of `x` in a trial.

  - The initial number of replications (`nrep`)
    can be smaller than `final_nrep`,
    such that the initial trials, though
    with lower precision (wider
    confidence intervals), are faster to
    run. As the solution is likely to be
    be found (values of `x` with estimated
    power close to the target value found),
    `nrep` will be increased successively
    to `final_nrep`, such that a trial is
    slower to run but has a higher precision.
    Other values that affect the speed,
    such as the number of values of `x`
    (`xs_per_trial`)
    and the number of iterations (`R`) in
    Monte Carlo confidence intervals
    and bootstrapping, are also increased
    successively.

- `x_final`

  - The value of `x` in the solution (e.g., with estimated power
    closest to the target value), if
    found.

- `power_final`, `nrep_final`, `ci_final`, `by_x_final`

  - Results based on `x_final`.

- Main functions used

  - `power4test_by_n()` and `power4test_by_es()`,
    for estimating the power levels for a set of values
    of `x`.

  - `power_curve()`, for estimating the relation between
    power and the value of `x`, based on the values of
    `x` having been examined.

  - The internal function `estimate_x_range()`, for
    determining the value(s) of `x` to be examined in a
    trial, given the value(s) examined so far and the
    tentative power curve.

# References