---
title: "Workflow of 'x_from_power()' Using the 'probabilistic_bisection' Algorithm"
output:
  html_document:
    fig.align: "center"
    toc: true
    number_sections: false
bibliography: references.bib
csl: apa.csl
---

```{r echo = FALSE}
library(DiagrammeR)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Goal

This technical appendix describes the
workflow of `x_from_power()` in
[power4mome](https://sfcheung.github.io/power4mome/)
when using the `"probabilistic_bisection"` algorithm,
based on the algorithm proposed by @waeber_bisection_2013,
applied to power analysis by @chalmers_solving_2024.

# `x_from_power()`

The following is the workflow of `x_from_power()`,
when the algorithm `"probabilistic_bisection"` is used.

```{r fig-n_from_power_pba, fig.cap = "From Power to x (Sample Size or Effect Size [Parameter Value])", echo = FALSE}
# https://mermaid.live/edit#pako:eNqtWP9P4kgU_1cmvWyCOd0FRUXuspeWgqKIG3VzOWBDBhigZ-mQdrrCGv_3e_O1M5XF_eFM1L55n_fevK8z7Ys3pTPiNb15TJ-nS5wy9BiOEoQ-fEDTGGdZSOZoRuY4jxmaR3Hc_K0NP53OH2UQSVOaJqBMwTqdEH7ewDIGNkgyU6gw5DhAcVzGtjFBURKxCMcSgOJosWSLlBALEVO6ttmTOCcFdx4lOI5-EBuxJTH4V2DIJmL7-IykK1DDyPhpLPTt2Q3X9Wug9QTvgHDQwK8MR14ngriwJUEbVEkQTRHJDkbetwOO8IPKyGtnLFrBrtAX-kzSJlrzf3VGMjaebMdJ5YALuYskq4AOqaIFKm5JuiAIxzFKSQZJzZoIUJtxTYOCGqC-ZoCBvW7IDPIB0fgO3lXmoD2hz1of_w1uXl5G3gONcxbRBM1pnsz-Gnmvr5LpV8rMJnogDDEq_MwUq4k2MoKHcjPqWXiiiWmkn5KUrPWzWlQ7Cm4C22BCWclo3__fTN2CpS_AxikRvtCcrXOGKjjhFlK6GgudiE7-JVMmYyZSPRh-Gnn3hOVpYgmOvE_fdC2go6PPPBGqK_LJIsXrpWmMIZg96kJSsPDygeB0ulTCIKWkJSWJlmRC2-1WybtlqJ3hWRbJcfRWRSohinoYMBqTNMFTgirwaGqsGpRwV-G9xOKEY5ezVAZm7Ei1SlIzEjNsy4kFLjOeF1IhSF2ShEdCJUEEO_oOXuTJVISnYuHbgG_RFYQbAAhqOieZ8VeHV9d7ZsQ6vMa7GaIJWJgL8ISXVQY9okvYFL1fvRxCL7c3YspJxaqBojlKyJRkGU630NffJPxKaefQhDwX8LsbS2n37R44-Gf7kP0us1_1VQYVGahEKbKlMqDIUIVWkW0VOUV2VESARH2qFi-V34q8Un5xzD8QYLna3V2BYpYPH1Me-R48q3oLRLnxswJ8Y5xrGigIeCh8OefRdEmmT8XEaflOinluN-MI5VmULGQ3RwhDXu51G7eCPQLCxJiLCSFJGlFefX8viep_nQC0wls0gYrtkTn7dM8H_QGq_BhHugwDXoVf1zNdszOSZBHbWhU7g0cL39nj76UqC77nHcUYXBW2AKI1XnGnxWmzTukET-CASsmC24bSEt1pog2Js3Id-Do9kgx0SuxcQ0yLRS3Y8lV-JBmq4NuULLmgLYMkiY6MgKP-sljT2q-kW4qQ5i_fAi4dPQp3FWhSOOdUp3UPQMMezhh6ggzgZAE6_kRmJv2lqxZ07ElW951kdYXwHjMFUvTAmqQsn4hTYDyFQcDQZ1S1QGKoLqM5ExYnWzA6BREcJQivOF7XQze0StK4LHcwpk8GNngfpsPgRFr0_u6wqpvT0AnZN4nnJu0kdn1UMEoWgoLjFmO3VeIYbaFkiJlUbNqwBz87NMXlcdjmf3n_8qllCqDr2L5WabX1qrWWQ4WGEvR1EWecbMcmXDoR1_uq7Nr336mza57HPtkwFOlbxCE_nNIc_j_DBZXPL4y-QEAMwJgeFJNZzkNhHuqPxjP6bHA9sYmKs330eP-1fYD8fogqtd_t3WdoBgebYvXh2jbL02IAuxYOjB-9wHKUUQpvGRQurnDPwzDYyPeI5pk85aUVcZkV99odYeu1-LF9T-N4gqdPJSUmDOIckDcjhBfQSfocB3l9Q3J9BkUdv_fQNoEJ-abfxsXaSSjTm2zdEMVwoNgo6X2yM0gWrv1LybYsmRvue3XU5x6bPhAnebRakVkEPsVbrebunUqAxljRlDi-KtbjMrJjH2XuRotKuBOxqNziTbTKV_LCkMGZhgE3O0B398jwxGmuOcVlCfrd7shrRfhydKjGcpvb961l3d49veiCb_3SCDQvhDD5nEYrZl9PqOCRUqqLGddzZ1yvVRISFad4YUnK7NRitLQe_y0vdE2Ffol2lbatobrL5f3jXoXNzMpBwdHpGfglC3oLpVjrnWpdZnXH4rV65zKgtkPeOorufFXWbnb6Ntbspe-XThElZyXoLlAlXFr3rWXXzO5jyXz3GL59LTVvce5N6sa6qWv3bqxbnwsE24aSb9suxi8WjbJAfQdQ8VQv6YZyQ3Mr1gcDeQ3zDr0VzAsczbym9yI-OHjg0QpOwSY8ik8oI2-UvAIQ54w-bJOp12RpTg69lOaLpdecwxQAKhdHaRhhCNRKQ9Y4GVBqSJhajKa38mOY-CYmIF7zxdt4zdpp_WOtcdo4r1VPTutn1cbxobf1mke16sX5x3q1Wj-unzfOjy8aJ6-H3g-h9vjjSaN-Vj-5OLk4rx6fVhuH3iLlvqgdgtskbfFbGKg_a9TOXv8Dd38JxQ
mermaid('
flowchart TD
  %% classDef default fill:#EEEEFF;
  %% classDef errornode fill:#FFDDDD;
  %% classDef startend fill:#DDFFDD;

  style initial fill: lightgreen
  style loop fill: lightblue
  style finalize fill: lightyellow
  style exit fill: lightyellow
  style terminate_k_final fill: lightgreen
  style exit_final fill: lightgreen
  style exit_pba fill: lightgreen

  ZA(["Find the x (n or es)"])
  AB("Estimate Power: power4test_by_n() or power4test_by_es()")
  AC("Merge all results: by_x_1")
  AB1("Use a fixed interval (for now)")

  BK{{"Solution found?"}}
  BKA("Solution found: Set to the solution: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")
  BKB("Solution not found: Set to NA: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")
  BM("Prepare the output (an x_from_power object)")

  ZZ[/"Return the output"/]

  ZA --> AB1

  subgraph initial [Pre-Iteration Search]

  AB1 --> AB
  AB --> AC

  end

  subgraph initialize [Prepare for the search]

  A0A("Set default tolernace (tol)")
  A0B("Set default HDR tolerance (hdr_power_tol)")
  A0C("Set default delta tolerance (delta_tol_f)")
  A0D("Generate the objective function (f)")
  A0E("Compute f values for the initial intervals")
  A0F{{"Is one of the bounds a solution?"}}
  A0G[["Extend the interval if necessary"]]
  A0H{{"Is the new interval OK?"}}
  A0I{{"Is one of the new bounds a solution?"}}

  AC --> A0A
  A0A --> A0B
  A0B --> A0C
  A0C --> A0D
  A0D --> A0E
  A0E --> A0F
  A0F -- No --> A0G
  A0G --> A0H
  A0H -- Yes --> A0I
  end

  subgraph loop [Trial Loop]

  BA("Start a trial")

  BB{{"A final check?"}}
  BCA("Compute f for x_i using nrep_i and R")
  BCB("Compute f for x_i using final_nrep and final_R")
  BD("Where the solution may be (Left/Right) (z_i)")
  BE("Update the density function (dfun_i)")
  BF{{"A final check?"}}
  BG{{"Is x_i a solution?"}}
  BH("Update x_i")
  BHB("Find probable region of power")

  A0I -- No --> BA

  BA --> BB

  BB -- Yes --> BCB
  BB -- No --> BCA
  BCA --> BD
  BCB --> BD
  BD --> BE
  BE --> BF
  BF -- Yes --> BG
  BF -- No --> BH
  BH --> BHB
  BG -- No --> BH
  BG -- Yes --> BHB
  BHB --> BHBA


  subgraph terminate_k [Last k changes < delta_tol?]

  BHBA{{"A final check?"}}
  BI{{"Is x_i a solution?"}}
  BIA{{"Last k changes < delta_tol?"}}
  BIB{{"pertubation_count > 0?"}}
  BIC("Shift x_i by a certain amount")
  BID("Update terminate_change_ok")
  BIZ("Update terminate_change_ok")

  BHBA -- Yes --> BI
    subgraph terminate_k_final [A final check]
      BI -- No --> BIA
      BIA -- Yes --> BIB
      BIB -- Yes --> BIC
      BIB -- No --> BID
    end
  BHBA -- No --> BIZ

  end

  subgraph exit [Exit the Loop?]

  BI -- Yes --> BJ
  BIA -- No --> BJ
  BIC --> BJ
  BID --> BJ

  BJ("Update any_terminate")
  BJA{{"A final check?"}}
  BJAA{{"Is x_i a solution?"}}
  BJZ("Next iteration, if run, will be a PBA iteration")
  BJZA("Start final_check_cooldown")
  BLA{{"(any_terminate TRUE) AND (1+ final checks done) AND (Not during final_check_cooldown)"}}
  BLB{{"Is x_i too close to a previous value checked in a final check?"}}
  BLC[["Rollback to a previous iteration and search again"]]
  BLCA("Set any_terminate to FALSE")
  BLD{{"any_terminate TRUE?"}}
  BLDA{{"Any final checks left?"}}
  BLDB{{"In final_check_cooldown?"}}
  BLE("Next iteration, if run, will be a final check")
  BMA{{"Is x_i a solution?"}}
  BNA("Exit the loop immediately")
  BOA{{"(any_terminate TRUE) AND (No more final checks) AND (This iteration is a final check)"}}
  BOB{{"(Maximum trials reached) OR (Maximum nrep reached)"}}

  BIZ --> BJ
  BJ --> BJA
  BJA -- Yes --> BJAA
  BJA -- No --> BLA
  BJAA -- Yes --> BMA
    subgraph exit_pba [A PBA iteration]
      BLA -- TRUE --> BLB
      BLB -- Yes --> BLC
      BLA -- FALSE --> BLD
      BLB -- No --> BLD
      BLC --> BLCA --> BLD
      BLD -- Yes --> BLDA -- Yes --> BLDB -- No --> BLE
    end

    subgraph exit_final [A final check]
      BJAA -- No --> BJZ
      BJZ --> BJZA
    end

  BLDB -- Yes --> BMA
  BLDA -- No --> BMA
  BLD -- No --> BMA
  BJZA --> BMA
  BLE --> BMA
  BMA -- No --> BOA
  BOA -- TRUE --> BNA
  BMA -- Yes --> BNA

  end

  BOA -- FALSE --> BOB
  BOB -- FALSE --> BA
  BOB -- TRUE --> BNA

  end

  subgraph finalize [Prepare the output]

  A0F -- Yes --> BK
  A0H -- No --> BK
  A0I -- Yes --> BK
  BNA --> BK

  BK -- Yes --> BKA
  BK -- No --> BKB
  BKA --> BM
  BKB --> BM

  end

  BM --> ZZ
', height = 3050, width = 800)
```

In `x_from_power`, `x` can be a sample
size (`n`) or a population value (`es`,
"effect size") of the selected model
parameter.

# Technical Details

The basic workflow is based on
the probabilistic bisection method
[waeber_bisection_2013]. In the current
implementation, the interval used in the
search is fixed
throughout the search. Users can set
the interval manually if so desired,
using the argument `x_interval`.

## Termination Criteria

The following criteria are used to decide
when the search will be ended.

- `tol`

    The tolerance used to determine whether
    the estimated power is "close enough"
    to the target power. If not set, determined
    by the confidence interval for a power
    equal to the target power, given `final_nrep`
    (the number of replications for the solution)
    and `ci_level` (the level of confidence
    for the estimated power): the distance
    from the confidence limit closet to
    the target power, multiplied by .90.
    Therefore, the larger the `final_nrep`,
    the smaller the tolerance. Used only
    when doing a final check using `final_nrep`.
    When calling `x_from_power()` and
    related functions, it can be set
    manually by the argument `tolerance`.
    It is used to determine whether a
    value of `x` is considered to be a
    solution. The default value ensured
    that the confidence interval of the
    solution includes the target power.

- `max_final_checks`

    The maximum number of final checks.
    It is possible that a candidate value
    turned out does not meet the
    tolerance (the estimated power not
    "close enough" to the target power).
    The search will then continue.
    The value `max_final_checks` determine
    the maximum number of final checks
    to try. If this number of checks is
    reached but solution not found, the
    search will terminate.
    The default value is 5, and
    can be changed through the
    `control` argument (e.g.,
    `control = list(max_final_checks = 2)`).

- `max_trials`

    The maximum number of trials (iterations)
    for the search. If this number of trials
    is reached, the search will stop.
    The default value is determined by
    the calling function, default to 100 for
    `probabilistic_bisection`, and
    can be changed through the argument
    `max_trials` of `x_from_power()` and
    friends.

- `total_nrep`

    The total number of replications
    for the whole search process. If this
    number is reached, the search will stop.
    The default value is determined by
    the calling function, default to 5000, and
    can be changed through the
    `control` argument (e.g.,
    `control = list(total_nrep = 2)`).

## Number of Replications in Each Iteration

The number of replications in a trial (`nrep_i`)
is `total_nrep / max_trials`.
Therefore, the default number of iterations
is 50. It will be adjusted upward such that
the estimated power in an iteration will not
be exactly equal to the target power. Therefore,
with a target power of .80, the default number
of iterations is 51, because an estimated power
of exactly .80 is possible with 50 replications.

## Final Checks

A value will be considered a solution
only if it passed the final check, carried
out using the number of replications equal
to `final_nrep`.

In an iteration, after the density function
is updated, the next iteration will be final
check if

- there are still unused attempts for
  final checks available (i.e., the number
  of attempted final checks is not equal to
  `max_final_checks`), and

- the iteration is not during the cooldown
  period after the previous final check,
  if any,

and at least one of the
following criteria is met (described
in details next):

- The range of changes of the candidate
  value (`x_i`), such as the sample size,
  in the `last_k` iterations is
  less than `delta_tol`.

- The range of changes of the function
  values (the deviation of the estimated
  power from the target power, if the goal
  is "close_enough") in the `last_k` iterations
  is less than `delta_tol_f`.

- The width of the (derived) highest-density region
  of the estimated power is less than or
  equal to `hdr_power_tol` *and* the
  region includes the target power.

If a solution is found in a final check,
the search will terminate.

If a solution is not found and this final
check is not the last one available,
the search will
resume. The next
`final_check_cooldown` iterations (default
to `last_k`) will be the cooldown period
and final check will not be carried out
regardless of other criteria.

## Final Check Criteria

The following criteria will be used to
determine whether a final check will be
conducted next.

- `hdr_power_tol`

    The tolerance used to determine whether
    the width of the highest-density region
    of power is narrow enough. If not set,
    to be conservative, it
    is set to the .90 of the width
    of the confidence interval for a power
    equal to the target power, given `final_nrep`.
    Therefore, the larger the `final_nrep`,
    the smaller the tolerance. It is used
    to determine whether a final check using
    `final_nrep` will be conducted next.
    To change it, set it through
    the `control` argument (e.g.,
    `control = list(hdr_power_tol = .02)`.

- `delta_tol`

    The tolerance used to determine whether
    the range of changes of `x` (sample size
    or effect size) in the last *k* trials
    is small enough. If the change is small
    enough, a solution may have been found.
    It is used
    to determine whether a final check using
    `final_nrep` will be conducted next.
    The default value is 2 for `n` (sample size)
    and .002 for `es` (effect size).
    To change it, set it through
    the `control` argument (e.g.,
    `control = list(delta_tol = .02)`.
    The number of trails (*k*), default to
    5, can be changed through the
    `control` argument (e.g.,
    `control = list(last_k = 3)`).

- `delta_tol_f`

    The tolerance used to determine whether
    the range of changes of the objective function
    value in the last *k* trials
    is small enough.  It is used
    to determine whether a final check using
    `final_nrep` will be conducted next.
    The default value is half of the width
    of the confidence interval at the target
    power.
    Although used in the check, this criterion
    is rarely met because the number of replications
    per trial (`nrep_i`) is designed to be small
    (e.g., around 50) for efficiency.
    To change it, set it through
    the `control` argument (e.g.,
    `control = list(delta_tol_f = .02)`.
    The number of trails (*k*), default to
    5, can be changed through the
    `control` argument (e.g.,
    `control = list(last_k = 3)`).

# Annotation

- `by_x_1`

  - The collection of all values tried and their results.
    It is updated whenever new value(s) is/are tried.

- `fit_1`

  - The latest power curve estimated by
    `power_curve`, using the values tried,
    stored in `by_x_1`. It is updated whenever
    `by_x_1` is updated.

- `x_i`

  - The value(s) for which power levels
    will be estimated in a trial.

- `final_nrep`

  - The desired number of replications
    for the solution. This value determines
    the desired level of precision
    (the width of the confidence interval)
    in the solution.

- `final_R`

  - The desired number of R for bootstrapping
    and Monte Carlo simulation
    for the solution.

- `x_final`

  - The value of `x` in the solution (e.g., with estimated power
    closest to the target value), if
    found.

- `power_final`, `nrep_final`, `ci_final`, `by_x_final`

  - Results based on `x_final`.

- Main functions used

  - `power4test_by_n()` and `power4test_by_es()`,
    for estimating the power levels for a set of values
    of `x`.

# References