---
title: "Workflow of 'x_from_power()' Using the 'bisection' Algorithm"
output:
  html_document:
    fig.align: "center"
    toc: true
    number_sections: false
bibliography: references.bib
csl: apa.csl
---

```{r echo = FALSE}
library(DiagrammeR)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Goal

This technical appendix describes the
workflow of `x_from_power()` in
[power4mome](https://sfcheung.github.io/power4mome/)
when using the `"bisection"` algorithm.

# `x_from_power()`

The following is the workflow of `x_from_power()`,
when the algorithm `"bisection"` is used.

```{r fig-n_from_power_bisection, fig.cap = "From Power to x (Sample Size or Effect Size [Parameter Value])", echo = FALSE}
# https://mermaid.live/edit#pako:eNqtVm1v2zYQ_isEhwIOkKSx4ziJBqywZDsoHHfF0qGY40CgbcrmJpEGRTV2g_z3HV9EU0qLfZmBINTdw3t7jke-4JVYUxzhLBfPqy2RCn0ZLThC796hVU7KckQztKYZqXKFMpbn0S9j-E0mv7ZBVEohORhzsMlkBL83sFKBD8rXDjUaaRygNK5Uh5wixpliJLcAlLPNVm0kpQEiF2IXqpd5RY_aarcmiqb7EHGgOeR3xIhK7Sr1AxcaMh92Hhd4wiBItaVojzocCYloebLATycaMRw-AmJEFZUF49TA6rAZB-k3WGzYN8obKklLKGMZoZKqdJ9Kwje0c6JtB4J0eUj3He3qybiKOws8LhUrICn0WTxTGaGd_tdXtFQaza2NppCW2oaNNgETMyo3FJE8iEI7SrsGZNNGZ2e_QW6OjGq5kWS39dE_fpb07CMkRxQTHD1QIlfbJwvWJbG7Y_cZ28_E6oFwuxgmRh5rXMPLTtLU8Grc3MPKm45NDRJJdQV0PcXyb7pSUF6UVXxloulkdbYarxNORAEU2w0ZAkIqWqIM6qQFS1HxdRnsGL28LPDHEglgU2QBBhFUirzSTj4s8OtrvWGsO2C8N61sOXa0swxxuqJlSeShJlFvmLz1YDt1_R-e7tzGcIf3Bn9s7eC-Wq7EiS-HE4x8tiBAn4QTj31OTjDxMTdwdy0uQaL1f0FdLaBFqaXzi9TN0yIU-HnQgwByVlp_ZMLoGtzVFBsKDYP7lAUbYlcfkP6khLHuh_GeKRsSKwq6ZlDH_BCYGQVlNgdRswRTS6F_EAxGEIBOCFQWcIg-oM49UAxgwmE85oqkSuQnoc9xYM_TpQ1wApPyGSx8ZWsFVfJ2YJSkNTJ91sqGwYntuB1pdZxtsrqylgrNhSunE8S-XA3O4qQhr7mOR74sP4E3Wige-6yb8ElDXsOToY83nrRFvn38HH_806w0wT5RpLfoCn_dEoWYLXNB1VasA-YNyszqWZXnVDrE8VgaQKwBZnqmq0rCVIGripWqhUo0KmYlNd3YUs51KDOyZ0VVIF4VS_ClD7nu7RImLllt6TGwYKMui4vN1WHY0r4NrEbGLaQPrgYkR2f-0NqS1Ii5F8VvRUkg8sJ5YyQ0xY02aY6K2Bmb2sHR7ConnPxIeBf6m7YmjLvG9ZWxI9LOi9-N7HgippqaBzcVYHzAqA1nw9TMooY6grtNwVE15up5EkH7ZYyT_NTem25t2XEfK1avOARUr53wOGqmceiSC9Vy-2n4_zmDp1d83wnfKnt4yNjXBFw8B5OjsYtsi3WqUt8vmU_8GOBJFDYjPDo0J_vUmztmOAOPISeOpw4MOUhGiiK1Lu1FHtzc02YLDEOxb4LYtnWzwaaup2f2K_Zf5mkzf3y_wH9QVUkeBLTA712fxDODn5tOx6e4gGIRtoZ38YtxhmFTQRc4gqV7DS_wgr8ClFRKPBz4CkdKVvQUS1FttjjK4ODDlx1iI0agYYsaAiN8LoT_hMtICTmz73DzHDcQHL3gPY66g_759cXtxc1Fr3s56F30-6f4gKOzq8HN-dXtZb_XHVwPbi4vb19P8XdjtXt-e3V9ddO77va615f9Qe8Ub6ROxgUIRaMyAUYVjga3r_8CHoO83g
mermaid('
flowchart TD
  %% classDef default fill:#EEEEFF;
  %% classDef errornode fill:#FFDDDD;
  %% classDef startend fill:#DDFFDD;

  style initial fill: lightgreen
  style loop fill: lightblue
  style update_x fill: lightyellow
  style output fill: lightgreen

  ZA(["Find the x (n or es)"])
  AA[["Determine the initial interval given the initial results: set_x_range() or set_x_range_by_x()"]]
  AB("Estimate Power: power4test_by_n() or power4test_by_es()")
  AC("Merge all results: by_x_1")

  ZA --> AA

  subgraph initial [Pre-Iteration Search]

    AA --> AB
    AB --> AC

  end

  AC --> BAB

  subgraph pre_loop [Pre-Loop]

    BAB("Create the objective function (f)")
    BAC("Compute the f values for the bounds")
    BAD{{"Is one of the bounds a solution?"}}
    BAE[["Extend the interval if necessary"]]
    BAF{{"Is one of the updated bounds a solution?"}}
    BAG{{"Is the updated interval valid?"}}

    BAB --> BAC
    BAC --> BAD
    BAD -- No --> BAE
    BAE --> BAF
    BAF -- No --> BAG

  end

  BAG -- Yes --> BA

  subgraph loop [Trial Loop]

    BA("Start a trial")
    BBA("Compute the function value for x_i")
    BBB{{"Is x_i a solution?"}}
    BBC("Exit loop immediately")
    BBD{{"Is the range of last k changes too small? (Less than delta_tol)"}}
    BBE{{"Is the interval too narrow? (Width less than min_interval_width)"}}
    BBF[["Expand the interval"]]

    BA --> BBA
    BBA --> BBB
    BBB -- Yes --> BBC
    BBB -- No --> BBD
    BBD -- Yes --> BBC
    BBD -- No --> BBE
    BBE -- Yes --> BBF
    BBE -- No --> BCA

    BBF --> BCA

    subgraph update_x [Update x_i]

      BCA{{"What is the method?"}}
      BCAA[["Muller method"]]
      BCAB[["power_curve assist"]]
      BCAC[["Bisection"]]
      BCAZ{{"Maximum number of trials reached?"}}

      BCA -- Muller --> BCAA
      BCA -- power_curve assist --> BCAB
      BCA -- Bisection --> BCAC

    end

    BCAA --> BCAZ
    BCAB --> BCAZ
    BCAC --> BCAZ

    BCAZ -- No --> BA
    BCAZ -- Yes --> BBC

  end

  BBC --> BK
  BAD -- Yes --> BK
  BAF -- Yes --> BK
  BAG -- No --> BK

  subgraph output [Prepare the Output]

    BK{{"Solution found?"}}
    BKA("Solution found: Set to the solution: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")
    BKB("Solution not found: Set to NA: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")
    %% BL("Determine x estimated by the power curve (used if solution not found): power_curve() --> x_estimated")
    BM("Prepare the output (an x_from_power object)")
    BK -- Yes --> BKA
    BK -- No --> BKB


  end

  BKA --> BM
  BKB --> BM

  ZZ[/"Return the output"/]

  BM --> ZZ


', height = 1980, width = 800)
```

In `x_from_power`, `x` can be a sample
size (`n`) or a population value (`es`,
"effect size") of the selected model
parameter.

# Annotation

- `by_x_1`

  - The collection of all values tried and their results.
    It is updated whenever new value(s) is/are tried.

- `fit_1`

  - The latest power curve estimated by
    `power_curve`, using the values tried,
    stored in `by_x_1`. It is updated whenever
    `by_x_1` is updated.

- `x_i`

  - The value for which power levels
    will be estimated in a trial.

- Termination criterion: `tol`

  - The tolerance used to determine whether
    the estimated power is a solution
    if the goal is "close enough"
    to the target power.

- Termination criterion: `delta_tol`

  - The tolerance used to determine whether
    the range of changes of `x` (sample size
    or effect size) in the last *k* trials
    is "too small." Unlike some other algorithm,
    having similar values across trials that are
    *not* solutions (e.g., not "close enough")
    indicate a problem in convergence,
    and so the algorithm will terminate.

- `x_final`

  - The value of `x` in the solution (e.g., with estimated power
    closest to the target value), if
    found.

- `power_final`, `nrep_final`, `ci_final`, `by_x_final`

  - Results based on `x_final`.

- Main functions used

  - `power4test_by_n()` and `power4test_by_es()`,
    for estimating the power levels for a set of values
    of `x`.

