---
title: "Workflow of 'x_from_power()' Using the 'probabilistic_bisection' Algorithm"
output:
  html_document:
    fig.align: "center"
    toc: true
    number_sections: false
bibliography: references.bib
csl: apa.csl
---

```{r echo = FALSE}
library(DiagrammeR)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Goal

This technical appendix describes the
workflow of `x_from_power()` in
[power4mome](https://sfcheung.github.io/power4mome/)
when using the `"probabilistic_bisection"` algorithm.

# `x_from_power()`

The following is the workflow of `x_from_power()`,
when the algorithm `"probabilistic_bisection"` is used.

```{r fig-n_from_power_pba, fig.cap = "From Power to x (Sample Size or Effect Size [Parameter Value])", echo = FALSE}
# https://mermaid.live/edit#pako:eNqtWP9P4kgU_1cmvWyCOd0FRUXuspeWgqKIG3VzOWBDBhigZ-mQdrrCGv_3e_O1M5XF_eFM1L55n_fevK8z7Ys3pTPiNb15TJ-nS5wy9BiOEoQ-fEDTGGdZSOZoRuY4jxmaR3Hc_K0NP53OH2UQSVOaJqBMwTqdEH7ewDIGNkgyU6gw5DhAcVzGtjFBURKxCMcSgOJosWSLlBALEVO6ttmTOCcFdx4lOI5-EBuxJTH4V2DIJmL7-IykK1DDyPhpLPTt2Q3X9Wug9QTvgHDQwK8MR14ngriwJUEbVEkQTRHJDkbetwOO8IPKyGtnLFrBrtAX-kzSJlrzf3VGMjaebMdJ5YALuYskq4AOqaIFKm5JuiAIxzFKSQZJzZoIUJtxTYOCGqC-ZoCBvW7IDPIB0fgO3lXmoD2hz1of_w1uXl5G3gONcxbRBM1pnsz-Gnmvr5LpV8rMJnogDDEq_MwUq4k2MoKHcjPqWXiiiWmkn5KUrPWzWlQ7Cm4C22BCWclo3__fTN2CpS_AxikRvtCcrXOGKjjhFlK6GgudiE7-JVMmYyZSPRh-Gnn3hOVpYgmOvE_fdC2go6PPPBGqK_LJIsXrpWmMIZg96kJSsPDygeB0ulTCIKWkJSWJlmRC2-1WybtlqJ3hWRbJcfRWRSohinoYMBqTNMFTgirwaGqsGpRwV-G9xOKEY5ezVAZm7Ei1SlIzEjNsy4kFLjOeF1IhSF2ShEdCJUEEO_oOXuTJVISnYuHbgG_RFYQbAAhqOieZ8VeHV9d7ZsQ6vMa7GaIJWJgL8ISXVQY9okvYFL1fvRxCL7c3YspJxaqBojlKyJRkGU630NffJPxKaefQhDwX8LsbS2n37R44-Gf7kP0us1_1VQYVGahEKbKlMqDIUIVWkW0VOUV2VESARH2qFi-V34q8Un5xzD8QYLna3V2BYpYPH1Me-R48q3oLRLnxswJ8Y5xrGigIeCh8OefRdEmmT8XEaflOinluN-MI5VmULGQ3RwhDXu51G7eCPQLCxJiLCSFJGlFefX8viep_nQC0wls0gYrtkTn7dM8H_QGq_BhHugwDXoVf1zNdszOSZBHbWhU7g0cL39nj76UqC77nHcUYXBW2AKI1XnGnxWmzTukET-CASsmC24bSEt1pog2Js3Id-Do9kgx0SuxcQ0yLRS3Y8lV-JBmq4NuULLmgLYMkiY6MgKP-sljT2q-kW4qQ5i_fAi4dPQp3FWhSOOdUp3UPQMMezhh6ggzgZAE6_kRmJv2lqxZ07ElW951kdYXwHjMFUvTAmqQsn4hTYDyFQcDQZ1S1QGKoLqM5ExYnWzA6BREcJQivOF7XQze0StK4LHcwpk8GNngfpsPgRFr0_u6wqpvT0AnZN4nnJu0kdn1UMEoWgoLjFmO3VeIYbaFkiJlUbNqwBz87NMXlcdjmf3n_8qllCqDr2L5WabX1qrWWQ4WGEvR1EWecbMcmXDoR1_uq7Nr336mza57HPtkwFOlbxCE_nNIc_j_DBZXPL4y-QEAMwJgeFJNZzkNhHuqPxjP6bHA9sYmKs330eP-1fYD8fogqtd_t3WdoBgebYvXh2jbL02IAuxYOjB-9wHKUUQpvGRQurnDPwzDYyPeI5pk85aUVcZkV99odYeu1-LF9T-N4gqdPJSUmDOIckDcjhBfQSfocB3l9Q3J9BkUdv_fQNoEJ-abfxsXaSSjTm2zdEMVwoNgo6X2yM0gWrv1LybYsmRvue3XU5x6bPhAnebRakVkEPsVbrebunUqAxljRlDi-KtbjMrJjH2XuRotKuBOxqNziTbTKV_LCkMGZhgE3O0B398jwxGmuOcVlCfrd7shrRfhydKjGcpvb961l3d49veiCb_3SCDQvhDD5nEYrZl9PqOCRUqqLGddzZ1yvVRISFad4YUnK7NRitLQe_y0vdE2Ffol2lbatobrL5f3jXoXNzMpBwdHpGfglC3oLpVjrnWpdZnXH4rV65zKgtkPeOorufFXWbnb6Ntbspe-XThElZyXoLlAlXFr3rWXXzO5jyXz3GL59LTVvce5N6sa6qWv3bqxbnwsE24aSb9suxi8WjbJAfQdQ8VQv6YZyQ3Mr1gcDeQ3zDr0VzAsczbym9yI-OHjg0QpOwSY8ik8oI2-UvAIQ54w-bJOp12RpTg69lOaLpdecwxQAKhdHaRhhCNRKQ9Y4GVBqSJhajKa38mOY-CYmIF7zxdt4zdpp_WOtcdo4r1VPTutn1cbxobf1mke16sX5x3q1Wj-unzfOjy8aJ6-H3g-h9vjjSaN-Vj-5OLk4rx6fVhuH3iLlvqgdgtskbfFbGKg_a9TOXv8Dd38JxQ
mermaid('
flowchart TD
  %% classDef default fill:#EEEEFF;
  %% classDef errornode fill:#FFDDDD;
  %% classDef startend fill:#DDFFDD;

  style initial fill: lightgreen
  style loop fill: lightblue
  style finalize fill: lightyellow
  style exit fill: lightyellow
  style terminate_k_final fill: lightgreen
  style exit_final fill: lightgreen
  style exit_pba fill: lightgreen

  ZA(["Find the x (n or es)"])
  AB("Estimate Power: power4test_by_n() or power4test_by_es()")
  AC("Merge all results: by_x_1")
  AB1("Use a fixed interval (for now)")

  BK{{"Solution found?"}}
  BKA("Solution found: Set to the solution: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")
  BKB("Solution not found: Set to NA: x_final, by_x_final, power_final, ci_final, nrep_final, i_final")
  BM("Prepare the output (an x_from_power object)")

  ZZ[/"Return the output"/]

  ZA --> AB1

  subgraph initial [Pre-Iteration Search]

  AB1 --> AB
  AB --> AC

  end

  subgraph initialize [Prepare for the search]

  A0A("Set default tolernace (tol)")
  A0B("Set default HDR tolerance (hdr_power_tol)")
  A0C("Set default delta tolerance (delta_tol_f)")
  A0D("Generate the objective function (f)")
  A0E("Compute f values for the initial intervals")
  A0F{{"Is one of the bounds a solution?"}}
  A0G[["Extend the interval if necessary"]]
  A0H{{"Is the new interval OK?"}}
  A0I{{"Is one of the new bounds a solution?"}}

  AC --> A0A
  A0A --> A0B
  A0B --> A0C
  A0C --> A0D
  A0D --> A0E
  A0E --> A0F
  A0F -- No --> A0G
  A0G --> A0H
  A0H -- Yes --> A0I
  end

  subgraph loop [Trial Loop]

  BA("Start a trial")

  BB{{"A final check?"}}
  BCA("Compute f for x_i using nrep_i and R")
  BCB("Compute f for x_i using final_nrep and final_R")
  BD("Where the solution may be (Left/Right) (z_i)")
  BE("Update the density function (dfun_i)")
  BF{{"A final check?"}}
  BG{{"Is x_i a solution?"}}
  BH("Update x_i")
  BHB("Find probable region of power")

  A0I -- No --> BA

  BA --> BB

  BB -- Yes --> BCB
  BB -- No --> BCA
  BCA --> BD
  BCB --> BD
  BD --> BE
  BE --> BF
  BF -- Yes --> BG
  BF -- No --> BH
  BH --> BHB
  BG -- No --> BH
  BG -- Yes --> BHB
  BHB --> BHBA


  subgraph terminate_k [Last k changes < delta_tol?]

  BHBA{{"A final check?"}}
  BI{{"Is x_i a solution?"}}
  BIA{{"Last k changes < delta_tol?"}}
  BIB{{"pertubation_count > 0?"}}
  BIC("Shift x_i by a certain amount")
  BID("Update terminate_change_ok")
  BIZ("Update terminate_change_ok")

  BHBA -- Yes --> BI
    subgraph terminate_k_final [A final check]
      BI -- No --> BIA
      BIA -- Yes --> BIB
      BIB -- Yes --> BIC
      BIB -- No --> BID
    end
  BHBA -- No --> BIZ

  end

  subgraph exit [Exit the Loop?]

  BI -- Yes --> BJ
  BIA -- No --> BJ
  BIC --> BJ
  BID --> BJ

  BJ("Update any_terminate")
  BJA{{"A final check?"}}
  BJAA{{"Is x_i a solution?"}}
  BJZ("Next iteration, if run, will be a PBA iteration")
  BJZA("Start final_check_cooldown")
  BLA{{"(any_terminate TRUE) AND (1+ final checks done) AND (Not during final_check_cooldown)"}}
  BLB{{"Is x_i too close to a previous value checked in a final check?"}}
  BLC[["Rollback to a previous iteration and search again"]]
  BLCA("Set any_terminate to FALSE")
  BLD{{"any_terminate TRUE?"}}
  BLDA{{"Any final checks left?"}}
  BLDB{{"In final_check_cooldown?"}}
  BLE("Next iteration, if run, will be a final check")
  BMA{{"Is x_i a solution?"}}
  BNA("Exit the loop immediately")
  BOA{{"(any_terminate TRUE) AND (No more final checks) AND (This iteration is a final check)"}}
  BOB{{"(Maximum trials reached) OR (Maximum nrep reached)"}}

  BIZ --> BJ
  BJ --> BJA
  BJA -- Yes --> BJAA
  BJA -- No --> BLA
  BJAA -- Yes --> BMA
    subgraph exit_pba [A PBA iteration]
      BLA -- TRUE --> BLB
      BLB -- Yes --> BLC
      BLA -- FALSE --> BLD
      BLB -- No --> BLD
      BLC --> BLCA --> BLD
      BLD -- Yes --> BLDA -- Yes --> BLDB -- No --> BLE
    end

    subgraph exit_final [A final check]
      BJAA -- No --> BJZ
      BJZ --> BJZA
    end

  BLDB -- Yes --> BMA
  BLDA -- No --> BMA
  BLD -- No --> BMA
  BJZA --> BMA
  BLE --> BMA
  BMA -- No --> BOA
  BOA -- TRUE --> BNA
  BMA -- Yes --> BNA

  end

  BOA -- FALSE --> BOB
  BOB -- FALSE --> BA
  BOB -- TRUE --> BNA

  end

  subgraph finalize [Prepare the output]

  A0F -- Yes --> BK
  A0H -- No --> BK
  A0I -- Yes --> BK
  BNA --> BK

  BK -- Yes --> BKA
  BK -- No --> BKB
  BKA --> BM
  BKB --> BM

  end

  BM --> ZZ
', height = 4050, width = 800)
```

In `x_from_power`, `x` can be a sample
size (`n`) or a population value (`es`,
"effect size") of the selected model
parameter.

# Annotation

- `by_x_1`

  - The collection of all values tried and their results.
    It is updated whenever new value(s) is/are tried.

- `fit_1`

  - The latest power curve estimated by
    `power_curve`, using the values tried,
    stored in `by_x_1`. It is updated whenever
    `by_x_1` is updated.

- `x_i`

  - The value(s) for which power levels
    will be estimated in a trial.

- `final_nrep`

  - The desired number of replications
    for the solution. This value determines
    the desired level of precision
    (the width of the confidence interval)
    in the solution.

- `final_R`

  - The desired number of R for bootstrapping
    and Monte Carlo simulation
    for the solution.

- Termination criterion: `tol`

  - The tolerance used to determine whether
    the estimated power is "close enough"
    to the target power. If not set, determined
    by the confidence interval for a power
    equal to the target power, given `final_nrep`.
    Therefore, the larger the `final_nrep`,
    the smaller the tolerance. Used only
    when doing a final check using `final_nrep`.

- Termination criterion: `hdr_power_tol`

  - The tolerance used to determine whether
    the width of the highest-density region
    of power is narrow enough. If not set, determined
    by the confidence interval for a power
    equal to the target power, given `final_nrep`.
    Therefore, the larger the `final_nrep`,
    the smaller the tolerance. It is used
    to determine whether a final check using
    `final_nrep` will be conducted next.

- Termination criterion: `delta_tol`

  - The tolerance used to determine whether
    the range of changes of `x` (sample size
    or effect size) in the last *k* trials
    is small enough.  It is used
    to determine whether a final check using
    `final_nrep` will be conducted next.

- Termination criterion: `delta_tol_f`

  - The tolerance used to determine whether
    the range of changes of the objective function
    value in the last *k* trials
    is small enough.  It is used
    to determine whether a final check using
    `final_nrep` will be conducted next.
    Although used in the check, this criterion
    is rarely met because the number of replications
    per trial (`nrep_i`) is designed to be small
    (e.g., around 50) for efficiency.

- `x_final`

  - The value of `x` in the solution (e.g., with estimated power
    closest to the target value), if
    found.

- `power_final`, `nrep_final`, `ci_final`, `by_x_final`

  - Results based on `x_final`.

- Main functions used

  - `power4test_by_n()` and `power4test_by_es()`,
    for estimating the power levels for a set of values
    of `x`.

